\maketitle  % command to print the title page with above variables
\setcounter{page}{1}
%---------------------------------------------------------------------
%                  영문 초록을 입력하시오
%---------------------------------------------------------------------
\begin{abstracts}     %this creates the heading for the abstract page
	\addcontentsline{toc}{section}{Abstract}  %%% TOC에 표시
	\noindent{
		Perovskite is attracting attention due to its superior performance, cheap price and easy manufacturing method compared to conventional ones in various photoelectric devices such as solar cell and LED. It was considered that there is a possibility of waveguiding effect due to the phenomenon that light is emitted from the outside of the crystal when a laser is fired in a typical perovskite material $\rm CsPbBr_3$ single crystal.
		
		This study analyzed the PL(Photoluminescence) data using the NT-MDT instrument to determine what causes the wave guiding effect and the process of it. From the analyzed PL data, we observed the tendency of PL data from the center to the edge of the single crystal.
		
		As a result, it was found that the waveguiding effect occurred, and the cause of the waveguiding effect was found from the peak tendency of exciton and biexciton.
		
		key word : perovskite, exciton, biexciton, waveguiding effect
	}
\end{abstracts}

\begin{abstractskor}
	페로브스카이트는 태양 전지, LED 등의 여러 광전소자 분야에서 기존의 것들에 비해 더 좋은 성능과 값싼 가격, 쉬운 제조 방법으로 인해 각광받고 있는 물질이다. 대표적인 페로브스카이트 물질인 $\rm CsPbBr_3$ 단결정에 레이저를 쏘았을 때에 결정의 바깥쪽에서 빛이 나오는 현상을 보고 그 원인으로 waveguiding effect의 가능성이 있다고 판단하였다. 
	
	본 연구는 $\rm CsPbBr_3$ 단결정에서 waveguiding effect가 일어나는지 확인하고 그 원인을 알아내기 위해서 NT-MDT 기기를 이용하여 PL(Photoluminescence) data를 얻고 분석하였다. 측정한 PL data는 Origin 9.0 프로그램을 이용하여 fitting 한 뒤, 단결정의 내부에서 가장자리로 가면서의 경기향성을 분석하였다. 
	
	이를 통해 전자가 에너지 준위가 더 높은 defect 쪽으로 많이 이동하게 되어 가장자리로 에너지가 모이는 waveguiding effect를 유발하여 결정의 가장자리가 밝게 빛나는 현상이 나타남을 확인할 수 있었다.
	
	키워드 : 페로브스카이트, exciton, biexciton, waveguiding effect
\end{abstractskor}
%----------------------------------------------
%   Table of Contents (자동 작성됨)
%----------------------------------------------
\cleardoublepage
\addcontentsline{toc}{section}{Contents}
\setcounter{secnumdepth}{3} % organisational level that receives a numbers
\setcounter{tocdepth}{3}    % print table of contents for level 3
\baselineskip=2.2em
\tableofcontents


%----------------------------------------------
%     List of Figures/Tables (자동 작성됨)
%----------------------------------------------
\cleardoublepage
\clearpage
\listoffigures	% 그림 목록과 캡션을 출력한다. 만약 논문에 그림이 없다면 이 줄의 맨 앞에 %기호를 넣어서 코멘트 처리한다.

\cleardoublepage
\clearpage
\listoftables  % 표 목록과 캡션을 출력한다. 만약 논문에 표가 없다면 이 줄의 맨 앞에 %기호를 넣어서 코멘트 처리한다.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Main Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\cleardoublepage
\clearpage
\renewcommand{\thepage}{\arabic{page}}
\setcounter{page}{1}



